---
title: "Team zobreth Project Proposal"
author: "Ethan Cook, Brian Betancourt, and Zongyu Li"
date: "2023-07-20"
output: html_document
---  
***  

Load the dataset:  
```{r}
vgsales = read.csv("vgsales.csv")
```

#### 1. The names of the students who will be contributing to the group project.
Zongyu Li, Ethan Cook, Brian Betancourt

#### 2. A tentative title for the project.
North American video game sales prediction

#### 3. Description of the data file 
(what they contain including number of variables and number of records). You do not necessarily have to list all the variables, but at least mention those of greatest importance.

This dataset contains a list of video games with sales greater than 100,000 copies. It was generated by a scrape of vgchartz.com.
Fields include:
* Rank - Ranking of overall sales
* Name - The games name
* Platform - Platform of the games release (i.e. PC,PS4, etc.)
* Year - Year of the game's release
* Genre - Genre of the game
* Publisher - Publisher of the game
* NA_Sales - Sales in North America (in millions)
* EU_Sales - Sales in Europe (in millions)
* JP_Sales - Sales in Japan (in millions)
* Other_Sales - Sales in the rest of the world (in millions)
* Global_Sales - Total worldwide sales.  

Number of observations: `r nrow(vgsales)`  
Number of variables: `r ncol(vgsales)`  


#### 4. Background information on the data sets, including specific citation of their source (so that I can also access it).  

https://www.kaggle.com/datasets/gregorut/videogamesales

The original datasource is from the website “vgchartz.com”, and has been extracted (or “scraped”) by the author of the dataset, as is described in the description on the Kaggle website. The script used by the author for scraping the data is also included here: GitHub - GregorUT/vgchartzScrape: Scrapy project for data capture of vgchartz.

#### 5. A brief statement of the business, science, research, or personal interest you have in the data set which you hope to explore.  

Scenario: A video game company doesn’t have a North American presence and is interested in releasing new games in the North American market. Given the historical data on video game sales from other companies, this company is interested in predicting its own video game sales. Then it can decide whether it should enter the North American market.  

#### 6. Evidence that the data can be loaded into `R`. Load the data, and print the first few values of the response variable as evidence.


```{r}
head(vgsales)$NA_Sales
```

***  

